---
title: "Extracting ICAO Carbon Calculator Data"
#output: bookdown::word_document2
output: bookdown::pdf_document2
date: "`r format(Sys.time(), '%d %B %Y')`"
bibliography: bib.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(pdftools)

icc_pdf <- "Methodology-ICAO-Carbon-Calculator_v11-2018.pdf"  # filename
pth     <- "./data-analytic/"                                 # where stored
pdf_fn  <- paste0(pth, icc_pdf)
```

# Overview

Environmental Sustainability has become a priority topic over the past months.
Within PBWG and the bi-lateral work, discussions revolve around establishing an initial "look-up" table for converting operational inefficiencies to associated fuel burn or CO2 estimates (or other emissions).

Publicly available documentation on fuel burn / emissions is limited.
ICAO maintains an ICAO Carbon Calculator that enables interested parties to inquire the fuel burn on an aerodrome-pair level.
The supporting methodology is published in [@icao_carbon_2018].


# Understanding Reading pdf files with R

reading in the pdf file

* pdf_text() generates a list of all pages of the pdf file

```{r}
icc_all <- pdftools::pdf_text(pdf_fn)
```

convert table page to a machine-readable table

```{r}
page_17 <- icc_all[17]  # extract page from list

page_17 <- page_17 %>%  
  read_lines() %>%      # use line breaks
  as_tibble()           # convert to tibble

page_17
```
we can skip the first rows

```{r}
skip_rows <- 4
page_17 <- page_17 %>% 
  filter(row_number() > skip_rows)

page_17
```

clean string and split columns

```{r}
# table has IATA Code and then ranges up-to 8500NM
col_names <- c("CODE","125","250","500","750","1000","1500","2000","2500", "3000", "3500","4000","4500","5000","5500","6000","6500","7000","7500","8000","8500")

page_17 %>% 
  mutate(
     value = trimws(value)
    ,value = gsub(pattern = "\\s+", replacement = "*", x = value)
    ) %>% 
  separate(
    col = value
    , into = col_names
    , fill = "right"
    )
```

# Extracting ICAO Carbon Calculator Fuel Burn Estimates

## Apendix C

wrap this into a function

```{r}
col_names <- c("CODE","125","250","500","750","1000","1500","2000","2500", "3000", "3500","4000","4500","5000","5500","6000","6500","7000","7500","8000","8500")

extract_icc_range_table <- function(.pdf_page, .skip_rows = 0, .col_names = col_names){
  range_table <- .pdf_page %>%
    read_lines() %>%      # use line breaks
    as_tibble() %>% 
    filter(row_number() > skip_rows) %>% 
    mutate(
       value = trimws(value)
      ,value = gsub(pattern = "\\s+", replacement = "*", x = value)
      ) %>% 
    separate(
      col = value
      , into = col_names
      , fill = "right"
      ) %>% 
    # filter all rows where each range column has no value
    filter(!if_all(col_names[2]:col_names[length(col_names)], is.na))
  return(range_table)
}
```

ICC document lists fuel consumption tables on page 17 through 23

```{r}
page_nbrs <-  17:23
fuel_stages <- page_nbrs %>% 
  purrr::map_dfr(.f = ~ icc_all[.x] %>% extract_icc_range_table) %>% 
  filter(CODE != "Code")
```

write out fuel_stages table for future use

write_csv(fuel_stages, "./data-analytic/ICC-FuelConsumption-v11-2018.csv")


## Apendix B - Aircraft Mapping

Follow same logic to read in mapping tables.
Note that there are multiple Aircraft-Equivalent mappings per page.

```{r}
extract_aircraft_mappings <- function(.pdf_page){
  mapping <- .pdf_page %>% 
    read_lines() %>% 
    as_tibble() %>% 
    filter(row_number() > row_number()[grepl("^Aircraft", value)]) %>% 
    mutate(
         value = trimws(value)
        ,value = gsub(pattern = "\\s+", replacement = "*", x = value)
    ) %>% separate(
        col = value
        , into = c("Aircraft","Equivalent","Aircraft_2","Equivalent_2","Aircraft_3","Equivalent_3","Aircraft_4","Equivalent_4")
        , fill = "right"
    )
  tmp1 <- mapping[,1:2]
  tmp2 <- mapping[,3:4]; names(tmp2) =  c("Aircraft","Equivalent")
  tmp3 <- mapping[,5:6]; names(tmp3) =  c("Aircraft","Equivalent")
  tmp4 <- mapping[,7:8]; names(tmp4) =  c("Aircraft","Equivalent")
  mapping <- bind_rows(tmp1, tmp2, tmp3, tmp4) %>% 
    # remove any incomplete combination / mapping (e.g. NAs, spurious page number)
    filter(!if_any(1:2, is.na)) 
  return(mapping)
}
```

```{r}
page_nbrs <-  14:16
ac_equi_map <- page_nbrs %>% 
  purrr::map_dfr(.f = ~ icc_all[.x] %>% extract_aircraft_mappings )
```

write_csv(fuel_stages, "./data-analytic/ICC-AircraftMapping-v11-2018.csv")

## ICAO Aircraft Type Designator to IATA Type Code Mapping

The ICC documentation uses - unfortunately - IATA Type Codes. 
Let's hack together a mapping table on top of our handwork for the ACERT look-up.

```{r}
# get something useful from wikipedia
library(rvest)

url <- "https://en.wikipedia.org/wiki/List_of_aircraft_type_designators"

wiki_html  <- read_html(url)
wiki_nodes <- wiki_html %>% 
  html_nodes(css = "table")

# only one table in nodes, but to be sure we fetch #1
wiki_tbl   <- wiki_nodes %>% 
  nth(1) %>% 
  html_table()

names(wiki_tbl) <- c("ICAO", "IATA", "MODEL")
```


```{r}
# another source
url2 <- "https://www.avcodes.co.uk/acrtypes.asp"

avcodes   <- read_html(url2)
avc_nodes <- avcodes %>% 
  html_nodes("table")
avc_tbl   <- avc_nodes %>% first() %>% html_table()

names(avc_tbl) <- c("IATA", "ICAO_AVCodes", "Manufacturer-Type-Model_AVCodes", "WTC")
```

```{r}
# another source
url3 <- "https://stringfixer.com/nl/List_of_aircraft_type_designators"

stringfix  <- read_html(url3)
sfix_nodes <- stringfix %>% html_nodes("table")
sfix_tbl   <- sfix_nodes %>% first() %>% html_table()

names(sfix_tbl) <- c("ICAO_sfix", "IATA_sfix", "MODEL_sfix")
```

combine all - for the time being

```{r}
rq <-  wiki_tbl %>% left_join(avc_tbl %>% rename(ICAO = ICAO_AVCodes)) %>% left_join(sfix_tbl %>% rename(ICAO = ICAO_sfix))
```
write_csv(rq, "./data-analytic/Aircraft-ICAO-IATA-Types.csv")


## Augment ICC Fuel Consumption

Appendix B recognises "equivalent" emitting aircraft types.
Thus, we have to augment the look up to account for these additional aircraft.

```{r}
fuel_stages2 <- fuel_stages %>% 
  mutate(VERSION = "ICC-v11-2018 (direct)")

which_equiv <- ac_equi_map %>% 
  mutate(CHECK = Aircraft != Equivalent) %>% 
  filter(CHECK == TRUE)

which_equiv <- which_equiv %>% select(-CHECK) %>% 
  inner_join(fuel_stages %>% rename(Equivalent = CODE)) %>% 
  mutate(VERSION = paste0("ICC-v11-2018 (equiv: ", Equivalent, ")")) %>% 
  rename(CODE = Aircraft) 

which_equiv <- which_equiv %>% 
  filter(CODE != "787")     # double accounting of 787

fuel_stages2 <- bind_rows(fuel_stages2, which_equiv) %>% 
  rename(IATA = CODE, EQUIVALENT = Equivalent) 
```

```{r}
fuel_stages2 <- fuel_stages2 %>% 
  left_join(avc_tbl %>% rename(ICAO = ICAO_AVCodes))
fuel_stages2 <- fuel_stages2 %>% select(ICAO, IATA, everything())
```

```{r}
missin_icao <- fuel_stages2 %>% 
  filter(ICAO == "" | is.na(ICAO))
```

fix by hand - check that we only have unique ICAO !

> missin_icao %>% left_join(sfix_tbl, by = c("IATA"="IATA_sfix")) %>%  select(ICAO_sfix, ICAO, IATA, everything()) %>% filter(!is.na(ICAO_sfix))
# A tibble: 5 x 28
  ICAO_sfix ICAO  IATA  `125` `250` `500` `750` `1000` `1500` `2000` `2500` `3000` `3500` `4000` `4500` `5000` `5500`
  <chr>     <chr> <chr> <chr> <chr> <chr> <chr> <chr>  <chr>  <chr>  <chr>  <chr>  <chr>  <chr>  <chr>  <chr>  <chr> 
1 A388      NA    388   5851  12016 17623 24940 32211  46695  61160  75638  90143  104681 119255 133865 148512 163196
2 A321      NA    32B   1909  3925  5270  7157  8970   12456  15818  19094  22308  NA     NA     NA     NA     NA    
3 A158      NA    A58   1543  3087  4064  5306  6478   NA     NA     NA     NA     NA     NA     NA     NA     NA    
4 E75L      NA    E7W   1113  2240  2989  3953  4890   6725   NA     NA     NA     NA     NA     NA     NA     NA    
5 E50P      NA    EP1   290   580   764   997   1218   NA     NA     NA     NA     NA     NA     NA     NA     NA   

write_csv(fuel_stages2, "./data-analytic/ICC-FuelConsumption-v11-2018-augmented.csv")


There exists already a A321, E75L.
Fixed by editing the csv (outside R).

# Conclusion

This document summarises the data preparatory steps for generating a look-up table for PBWG and the bi-regional comparison work.
It is based on the extraction of fuel burn estimates from the ICAO Carbon Calculator Methodology document, v11 2018.

The data has been extracted from the respective Appendix C.
With Appendix B, the lookup has been expanded for a set of aircraft types that show - in accordance with the methodology document - similar fuel burn characteristics. 

The comparison work is based on ICAO Aircraft Type Designators [@icao_doc8643_2022].

== initial lookup table!

# References {-}
